RickView: Lightweight Standalone Knowledge Graph Browsing

### 2024-03-21 paper.tex
more D2R2 paper

%
\begin{abstract}
The standard backends for RDF browsers are SPARQL endpoints, which allow expressive queries, but they are overengineered for the simple task of browsing.
RickView is an URI dereferenciation tool and RDF browser that follows an alternative approach of directly querying an in-memory dataset bypassing SPARQL.
It is implemented in Rust, which enables a high level of performance comparable to C and C++ while being memory-safe and thread-safe by design.
RickView is published as a single binary, optionally wrapped in a compact container image, without dependencies on external components.
It is adaptable to a given knowledge base via defaults that can be changed in a configuration file and overridden with environment variables.
It supports the binary RDF format Header Dictionary Triples (HDT), which allows querying compressed RDF in memory.
These features enable the stable long-term, low-maintenance, robust, simple and fast publication of multiple RDF knowledge graphs on low-end hardware.
\end{abstract}
\begin{keywords}Linked Data browser \sep RDF visualization.\end{keywords}

@@

\section{Introduction}\label{summary}
% integrate this, translated from German
Symbolic AI methods involve formal knowledge representation in the form of ontologies and knowledge graphs, which are described using the Web Ontology Language (OWL) and the Resource Description Framework (RDF). In the vision of the Semantic Web, existing data silos are replaced by providing represented knowledge in its entirety in forms that are accessible to both humans and agents (programs).
These methods are widely used by industry, for example, for intelligent searches through the Google Knowledge Graph, but contrary to the vision, both the knowledge graphs and the tools are not shared with the public.
On the other hand, a multitude of knowledge graphs are created in academic research projects with limited duration, which lack the personnel or technical resources to build and operate the necessary technical infrastructure. Even if successful, in the event of a critical component failure, responsibility is unclear as those responsible are either working on new projects or have left the institution.
The result is knowledge graphs that either are not published according to established standards, or require maintenance, energy, and resource-intensive operations, ultimately leading to longer or even permanent outages.

@@

\section{Use Cases}
Current RickView deployments include the following knowledge graphs:

\begin{table}
\begin{tabular}{lrSSSSS}
\toprule
Knowledge Graph	&\textnormal{Size in MiB}	&\textnormal{LodView}	&\textnormal{Virtuoso 7.2.12}	&\textnormal{LodView+Virtuoso}	&\textnormal{RickView 0.2.13}\\
\midrule
SNIK v24.03		&2.0						&						&127.8							&								&23.4\\
HITO v24.02		&1.0						&						&110.2							&								&18.2\\
ANNO			&\\
LinkedSpending	&\\
\bottomrule
\end{tabular}
\label{tab:stats}
% todo correct versions
\caption{Knowledge graph serialized size (uncompressed RDF Turtle) and total deployment RAM consumption of LodView 1.2.3, Virtuoso 1.2.3 and RickView 1.2.3.
File sizes and memory usage in MiB.}
\end{table}

\subsection{Small Knowledge Bases: SNIK, HITO and ANNO}
SNIK\footnote{The \emph{Semantic Network of Information Management in Hospitals}, deployed at \url{https://www.snik.eu/ontology}.}~\citep{snik},
%HITO\footnote{The \emph{Health IT Ontology}, deployed at \url{https://hitontology.eu/ontology}.}~\citep{hito}
HITO\footnote{The Health IT Ontology [HITO] for systematic description of application systems and software products in healthcare.}~\citep{hito}
and ANNO\footnote{The \emph{Anthropological Notation Ontology}, deployed at \url{https://annosaxfdm.de/ontology}.}~\citep{anno}
exemplify small knowledge bases that were threatening the stablity of the server they were deployed on, together with many other small-scale services, due to RAM consumption, which led to the initial development of RickView.
In those cases, the overhead of multiple docker containers with a SPARQL endpoint, Java virtual machine and LodView, is disproportionate to the size of the knowledge bases, see \cref{tab:stats}.
%SNIK, HITO and many other small-scale services were threatening the stability of the .

@@

\subsection{HDT}

>>>>

% clarify relationship to resilience
\begin{abstract}
The standard backends for RDF browsers are SPARQL endpoints, which allow expressive queries, but they are overengineered for the simple task of browsing.
RickView is an URI dereferenciation tool and RDF browser that follows an alternative approach of directly querying an in-memory dataset bypassing SPARQL.
It is implemented in Rust, which enables a high level of performance comparable to C and C++ while being memory-safe and thread-safe by design.
RickView is published as a single binary, optionally wrapped in a compact container image, without dependencies on external components.
It is adaptable to a given knowledge base via defaults that can be changed in a configuration file and overridden with environment variables.
It supports the binary RDF format Header Dictionary Triples (HDT), which allows querying compressed RDF in memory.
These features enable the stable long-term, low-maintenance, robust, simple and fast publication of multiple RDF knowledge graphs on low-end hardware.
\end{abstract}
\begin{keywords}Linked Data browser \sep RDF visualization.\end{keywords}

@@

\section{Introduction}\label{summary}
% integrate this, translated from German
% D2R2’24: Third International Workshop on Linked Data-driven Resilience Research 2024
% We can skip this explanation as the target audience knows it already.
%Symbolic AI methods involve formal knowledge representation in the form of ontologies and knowledge graphs, which are described using the Web Ontology Language (OWL) and the Resource Description Framework (RDF).
%In the vision of the Semantic Web, existing data silos are replaced by providing represented knowledge in its entirety in forms that are accessible to both humans and agents (programs).
These methods are widely used by industry, for example, for intelligent searches through the Google Knowledge Graph, but contrary to the vision, both the knowledge graphs and the tools are not shared with the public.
On the other hand, a multitude of knowledge graphs are created in academic research projects with limited duration, which lack the personnel or technical resources to build and operate the necessary technical infrastructure. Even if successful, in the event of a critical component failure, responsibility is unclear as those responsible are either working on new projects or have left the institution.
The result is knowledge graphs that either are not published according to established standards, or require maintenance, energy, and resource-intensive operations, ultimately leading to longer or even permanent outages.

@@

\subsection{RDF Libraries}

\subsection{Header Dictionary Triples (HDT)}

% ******************************************************************
\section{Use Cases}
Current RickView deployments include the following knowledge graphs:

\begin{table}
\begin{tabular}{lSSSSSSS}
\toprule
% triples SNIK 44584, HITO, ANNO, linkedspending					
Knowledge Graph	&\textnormal{k triples}		&\textnormal{Size in MiB}	&\textnormal{LodView}	&\textnormal{Virtuoso}	&\textnormal{LodView+Virtuoso}	&\textnormal{RickView}\\
\midrule
SNIK v24.03		&44.6						&2.0						&						&127.8							&								&23.4\\
HITO v24.02		&							&1.0						&						&110.2							&								&18.2\\
ANNO			&							&\\
LinkedSpending	&							&28256.6					&						&								&								&1973.2\\
\bottomrule
\end{tabular}
\label{tab:stats}
% todo correct versions
\caption{Knowledge graph serialized size (uncompressed N-Triples) and total deployment RAM consumption of LodView 1.2.3, Virtuoso 7.2.12 and RickView 0.2.13 of SNIK v24.03, HITO v24.02, ANNO v00000 and LinkedSpending v2015.
LinkedSpending uses the compressed HDT mode, see \cref{sec:hdt}.
File sizes and memory usage in MiB rounded to one decimal place.}
\end{table}

\subsection{Small Knowledge Bases: SNIK, HITO and ANNO}
SNIK\footnote{The \emph{Semantic Network of Information Management in Hospitals}, deployed at \url{https://www.snik.eu/ontology}.}~\citep{snik},
%HITO\footnote{The \emph{Health IT Ontology}, deployed at \url{https://hitontology.eu/ontology}.}~\citep{hito}
HITO\footnote{The Health IT Ontology [HITO] for systematic description of application systems and software products in healthcare.}~\citep{hito}
and ANNO\footnote{The \emph{Anthropological Notation Ontology}, deployed at \url{https://annosaxfdm.de/ontology}.}~\citep{anno}
exemplify small knowledge bases that were threatening the stablity of the server they were deployed on, together with many other small-scale services, due to RAM consumption, which led to the initial development of RickView.
In those cases, the overhead of multiple Docker containers with a SPARQL endpoint, Java virtual machine and LodView, is disproportionate to the size of the knowledge bases, see \cref{tab:stats}.
This overhead can be reduced by sharing a single SPARQL endpoint, Java virtual machine (if required) and web server for all knowledge graphs published by a single server, but that is not practical for a modern, container based deployment,
which must be isolated from others in case of failure and must be able to be developed, updated and moved separately.
%SNIK, HITO and many other small-scale services were threatening the stability of the .

@@

\subsection{HDT}\label{sec:hdt}
